Learning to Optimize Real-World Programs

### 2021-02-04 paper.tex
1:17 pm update

% inserted by alex
\usepackage{amsmath,amssymb}
\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}
\DeclareMathOperator{\E}{\mathbb{E}}
\usepackage[dvipsnames]{xcolor}
\newcommand{\GH}{\textsc{Github}}
\usepackage{listings}
\lstdefinelanguage
  [x64]{Assembler}
  [x86masm]{Assembler}
  {basicstyle=\ttfamily\bfseries\scriptsize,
  frame=single,
  keepspaces=true,
  framesep=5pt,
  xleftmargin=10pt}
\lstset{language=[x64]{Assembler}}

@@

Program optimization is a classical problem in computer science that has existed for over 50 years \cite{mckeeman1965peephole, allen1971}, and most years, entire tracks at compiler conferences are dedicated to it. \ed{I think there is a sentence missing here that explains most optimizing compilers use hand-written/heuristics-based/rule-based optimizations.}  Beyond heuristic-based and semantics-preserving optimizations, automated methods based on brute-force search, heuristic search, and SMT solvers have been shown to outperform compiler-based optimizations. However, in practice they are generally \ed{too?} costly to use for all code during the compilation process.

Deep reinforcement learning has shown promise in addressing challenging tasks that search-based techniques had once struggled to scale to \cite{silver2017mastering}. Similar methods grounded in machine-learning \gn{don't hyphenate noun phrases unless they're used as adjectives} could hold promise for speeding up automated program optimization by applying learned knowledge quickly, as opposed to running a costly search or solver algorithm at compile-time \gn{same}. However, to our knowledge, current research in machine-learning-based program optimization has been generally constrained to the problem of expression simplification in a relatively simple domain-specific language (DSL) on synthetically created expressions and testcases \cite{shi2020}. Similarly, much work in machine-learning-based program synthesis  has been performed on synthetically-created datasets \cite{parisotto2016neuro,bunel2018leveraging}. While work on these datasets is an important research direction, these methods may have not yet been evaluated on realistic programs and input/output (I/O) examples, such as those found in open source code projects. Additionally, many of these machine-learning-based models often require access to a grammar and may not be directly extendable to program representations in-which a grammar may not be available or easy to construct.

\gn{Introducing a dataset doesn't sound exciting as discussing the core experimental result, which I think you should do first. Like ``we demonstrate that ML-based techniques can optimize real-world programs''}
In this work, we introduce a dataset for program optimization and program synthesis mined from real-world projects on \textsc{Github}. It consists of functions compiled to x86-64 assembly and automatically generated testcases with comprehensive coverage. \ed{What kind of coverage?  More importantly: I don't think we actually get comprehensive coverage in practice}.  If we want to keep this statement, we need to  In order to avoid learning spurious optimizations on a lack of testcase coverage, we utilize an SMT solver to prove equivalence of the optimizations. Based on our preliminary results, we argue that future machine-learning-based program synthesis work should strive to incorporate solvers and/or exhaustive comprehensive testcase coverage for evaluation. Lastly, we also show that an iterative hill-climbing algorithm, a special case of  regularized off-policy reinforcement learning (RL), is capable of discovering new optimizations. \pcyin{Maybe one or two more sentences to briefly discuss results?}

@@

The general undecidability of program equivalence means that there may always be room for improvement in optimizing programs~\citep{rice1953}. This is especially true as hardware options and performance goals become more diverse: what transformations are best for a scenario may vary greatly on certain hardware or depending on performance objectives such as such as energy consumption or runtime. \alex{citation here ?}

Given these challenges and the time-consuming nature of expert-written optimizations, automated methods of program optimization can be desirable.
\ed{This language is a bit ambiguous.  Modern optimizing compilers are automated even though they are based on manually specified expert rules.}

There are classes of functions that are amenable to ``superoptimization", where the optimal sequence of calculations is found \citep{massalin1987superoptimizer}. However, these are limited to very small sequences of machine instructions. Moreover, there are other classes of optimizations that are difficult or impossible to express in universal heuristics and semantics-preserving transformations, especially for program optimizations that might require domain-specific knowledge \citep{rinard2006}, are written by hand (e.g.,~OpenBLAS\footnote{https://www.openblas.net}) or adapted to specific hardware \citep{fftw}.

State of the art methods of automatic program optimization either rely on search-based procedures \citep{schkufza2013stochastic}, or constraint-based methods \citep{sasnauskas2017}. However, these methods have difficulty scaling to larger problems, and as a result, typically do not meet the performance requirements of real development scenarios at compile-time.
\ed{Are these super-optimizers too?} \alex{Yes in the sense I believe both claim to be superoptimizers; perhaps ``no" if we are strict in defining a superoptimzier to find the most optimzal sequence. I am not sure how souper works (it uses SMT solvers I think) but I know that STOKE (Schkufza 2013) cannot guaranteed optimal. I think we can call these ``superoptimziers" as that does seem to be the norm in the literature. }

@@

\paragraph{Machine-Learning-Based Program Synthesis} Program synthesis is the task of automatically generating a program that is consistent with a certain specification. Recent work in machine-learning-based program synthesis has traditionally used I/O examples as a way to provide the program specification. In \cite{bunel2018leveraging} the authors perform experiments in the Karel programming language \cite{pattis1981karel} an educational programming language that manipulates the actions of a robot in a gridworld.
% The programming language contains control flow operations such as loops and conditionals. 
In these tasks, programs are synthetically generated by sampling from a domain-specific language (DSL) \alex{This is the second time I abbreviate DSL, I now do it also in the abstract}. I/O examples are randomly generated, but with full-branch coverage \ed{Seems like a contradiction?}, meaning each individual conditional must be triggered by at least one  testcase; however, this does not imply comprehensive test coverage of all paths through program execution. While the authors RL experiments achieved highest performance with top-1 program synthesis accuracy; maximum-likelihood estimation (MLE) on target programs gradually outperformed RL as the beam-size increased to 50.

@@

In describing the task of neural program optimization, we build on notation from the program synthesis formulation in \cite{bunel2018leveraging} \gn{use ``citet'' for citations that are part of a sentence. Also in other places (like the future work)}. The task of program optimization through neural sequence models 
\gn{Do we need ``through neural sequence models''? Be best to skip this and make it more broad in the problem formulation.} 
is transducing an (unoptimized) program $\text{F}_u$ into an optimized one $\text{F}_o$, where $\text{F}_o$ is better than $\text{F}_u$ under some cost function $\mathbfcal{C}$, such as runtime, memory footprint or energy consumption.
%and equivalent under a distance metric $\mathbfcal{D}$. 
%Our goal is to learn a neural model  $f_{\theta}: \text{F}_u \mapsto \text{F}_o$. 
%\pcyin{Defer introducing $\mathcal{D}$ to the later paragraph where this notation is used.}

Because each program is a function that maps inputs to outputs, we will use $I$ to represent the hardware state prior to executing the program (i.e., input) and $O$ to represent the hardware state after executing the program (i.e., output).

Specifically, our goal is to learn a neural \gn{again, is ``neural'' necessary?} model $f_{\theta}: \text{F}_u \mapsto \text{F}_o$ such that a model-generated (optimized) program $\hat{F_o}$ attains lower cost under the cost function $\mathcal{C}$ evaluated on a suite of $K$ input-output examples $\{ IO \}_{k=1}^K$:
%on our synthesized program $\hat{\text{F}}_o$,
%and observing $\mathcal{C}$ which may measure approximate runtime, energy consumption, or other goals. Furthermore, we may evaluate our distance function $\mathcal{D}$ either as equivalence under all testcases, or through verification by using an SMT solver. We consider a program optimized when: 
% cost function $\mathbfcal{C}$ which is able to evaluate desired program behavior, such as approximate runtime or energy consumption. For the transductive objective, the learned model's objective is then the following: 
\begin{equation}
    \label{eqn:optimizaiton_goal}
    \begin{split}
        \mathbfcal{C} \Big(\hat{\textrm{F}}_{o}; \{IO_k\}_{k=1...K} \Big)  \ 
        < \ 
        \mathbfcal{C} \Big(\textrm{F}_{u}; \{IO_k\}_{k=1...K}\Big) \\
         s.t. \;\; \mathbfcal{D}( \ 
                        \hat{\textrm{F}}_{o}, \textrm{F}_{u} ) \ 
                        = 0  % \;\; \
        %  \hat{\textrm{F}}_{opt}^i(I_k^i) = \ 
        %         O_k^i , \\
            % \forall  \; k \;  \in 1...K 
    \end{split}
\end{equation}

where $\mathcal{D}(\cdot)$ is a distance metric that measures if the model optimized program $\hat{F}_o$ is functionally equivalent to the original one $F_u$. For conciseness, we will abbreviate $\mathbfcal{C} (\hat{\textrm{F}}_{o}; \{IO_k\}_{k=1...K})$  as $\mathbfcal{C} (\hat{\textrm{F}}_{o})$.

In situations in which we desire to learn optimizations through demonstrations, we could collect ground-truth optimized programs $\textrm{F}_o$ from an optimizing compiler.
For example, as we elaborate in Section xxx, the C program compiler GCC could be configured to perform different levels of optimization.

In situations in which we desire to learn optimizations through demonstrations, we will denote a ground-truth optimized program, for example, from an optimizing compiler, as $\textrm{F}_o$. In our later experiments, prior to training, we initialize a dataset $D_o$ of $N$ tuples of an I/O test suite, an unoptimized function, and compiler-optimized function as demonstration:

\begin{equation} 
    \begin{split}
    \label{eqn:init_dataset}
        D_o = \
                \bigg\{
                    \Big( \
                        \{IO_k^i\}_{k=1...K}, \
                        \textrm{F}_{u}^i, \
                        \textrm{F}_{o}^i, \
                        % \textit{LiveOut}^i\
                    \Big) \
                \bigg\}_{i = 1...N} \\
        % s.t. \;\;   \textrm{F}_{opt}^i(I_k^i) = \ 
        %     \textrm{F}_{uopt}^i(I_k^i) = O_k^i  \\
        % \forall  \; k \;  \in 1...K \;\;  \ 
        %     \forall \; i \;  \in 1...N 
        % C(\textrm{F}_{uopt}^i; \{IO_k^i\}_{k=1...K})  \ 
        %     \leq \ 
        %     C(\textrm{F}_{uopt}^i; \{IO_k^i\}_{k=1...K}) 
    \end{split}
\end{equation}

In \S\ref{sec:benchmark} we describe one concrete instantiation of this problem formulation on real-world programs, and in \S\ref{sec:learning} we describe our methodology for modeling and learning $f_\theta$ within this setup.

@@

@@

In order to create a dataset of functions whose re-writes could be executed and evaluated, we further mined a subset of 15,175 functions for training, 775 for validation, and 942 for testing. \ed{The jump from 1.77 million examples to ~17K deserves more of an explanation.} \gn{Right, and it's probably because only about 17k finished processing right? It'd be nice to change 1.77 million to the number of ones we actually \emph{tried} to process, and explain why some were filtered out.} We were able to create testcases through random generation using the \textsc{STOKE} toolkit\footnote{\url{https://github.com/StanfordPL/stoke}}. \gn{Footnotes should come after punctuation.}  In order to create conservative benchmarks \gn{Needs more explanation.}, we eliminated any functions in our dataset that caused segmentation faults either when testcases were executed on the original unoptimized or optimized function. \ed{Does this mean that if a function dereferences a pointer argument it will likely be excluded?} We further removed any functions where the compiler optimized program was not logically equivalent to the unoptimized program using an SMT solver and equivalent under all testcases. \ed{I think this also needs more explanation.  It makes it sound like the compiler is to blame, but that's not really the case.} To further avoid learning spurious optimizations, we required that our programs were not equivalent under testcases to three pathological spurious program examples. 
%\pyin{Could you explain the last sentence more?}
\ed{I also found the previous sentence confusing.}

@@

Our neural optimizer \gn{not optimizer, model} uses a standard Transformer-based encoder \citep{vaswani2017attention}, but swapped our decoder with an LSTM-based decoder \citep{hochreiter1997long,chen-etal-2018-best}.
This was mainly done to speed-up prediction within the iterative training loop described in \S\ref{sec:hillclimbing}, as transformers suffer from quadratic decoding time in sequence length \citep{chen-etal-2018-best}.
\gn{More detail. What are your inputs and outputs. Do you do subword tokenization? Can you give an example of what the inputs and outputs look like?}

\gn{You have a pretty extensive description below on why you \emph{don't} do what you don't do, but only one sentence explaining what you actually do :)}
While it is often advantageous to leverage a grammar for code generation tasks \citep{parisotto2016neuro,yin-neubig-2017-syntactic}, we chose not to do so in our experiments. A primary reason was for simplicity. An additional reason was access to large amounts of data for supervised learning, which may implicitly learn syntax. Previous work in machine-learning-based program synthesis has demonstrated that while a grammar may help greatly in low-data program synthesis regimes, it may do little to improve both MLE and RL-based experiments when extensive data is available for pre-training \cite{bunel2018leveraging}.

@@

\pcyin{When explaining a model/algorithm, it's important to first briefly describe the intuition behind each component/sub-section at the very beginning. Can we say something like the follows here: With the model pre-trained on compiler-optimized examples, we then aim to learn more efficient optimization strategies by directly minimizing the expected cost of model-optimized programs.}
Policy gradient algorithms seem like a natural fit for learning how to optimize programs; however, methods based on maximum-likelihood estimation, or imitation learning, provide advantages in terms of stability. Training large neural networks with policy gradient algorithms can be difficult to tune. These difficulties are often attributed to the gradient's high variance which is exacerbated by other issues in program synthesis: a general sparsity of rewards, credit assignment over long sequences, and brittleness of program syntax and semantics. \alex{I think this point here contradicts my earlier point about needing to sample from a grammar.... If a grammar really isn't that necessary after pre-training, then is brittle syntax really an issue?? It may be good to leave this out or accept either the nuance proposed / contradiction}
\pcyin{Perhaps just removing ``brittleness of program syntax'' is good?}

@@

\begin{algorithm}
\label{alg:hill_climbing}
\begin{algorithmic}
\STATE Initialize model $\textbf{Model}$ parameters $\theta$ from pre-trained model \\
\STATE Initialize dataset of program function pairs and testcases: \\  
    $D = D_o =  \Big\{ \big( \
    \{IO_k^i\}_{k=1...K}, \
    \textrm{F}_{u}^i, \
    \textrm{F}_{o}^i, \
    % \textit{LiveOut}^i\
    \big) \ 
    \Big\}_{i=1...N}$ \\
 \WHILE{not converged}
  \STATE Sample a batch $B_{\mathrm{ex}}$ from $D_o$\\
  \FOR{ $\big( \
    \{IO_k^i\}_{k=1...K}, \
    \textrm{F}_{u}^i, \
    \textrm{F}_{o}^i, \
    % \textit{LiveOut}^i\
    \big)$  in $B_{\mathrm{ex}}$} 
    \STATE  sample $\hat{\textrm{F}}_{o}^i \; \sim \; f_{\theta}(\textrm{F}_{u}^i)$ \\
    \STATE calculate 
    $\mathbfcal{C}(\textrm{F}_{o}^i)$, 
    $\mathbfcal{C}(\hat{\textrm{F}}_{o}^i)$, \ 
    and $\mathbfcal{D} (\hat{\textrm{F}}_{o}^i, \textrm{F}_{o}^i)$ 
    \IF{$\mathbfcal{D} (\hat{\textrm{F}}_{o}^i, \textrm{F}_{o}^i) == 0$ and \ 
    $\mathbfcal{C}(\hat{\textrm{F}}_{o}^i) < \ 
    \mathbfcal{C}(\textrm{F}_{o}^i) $ }
        \STATE  Replace $\textrm{F}_{o}^i$ with sample $\hat{\textrm{F}}_{o}^i$ in $D$ \\
    \ENDIF
  \ENDFOR
  \STATE Sample a batch $B_{\mathrm{tr}}$ from $D$\\
  \STATE Update $\theta$ via supervised learning on $B_{\mathrm{tr}}$ from $D$ \\
 \ENDWHILE\caption{Hill-Climbing for Program Optimization}
\end{algorithmic}
\end{algorithm}

@@

\paragraph{Canonicalizing Programs} The x86-64 assembly was to canonicalize locations within programs and pre-processed  with byte-pair encoding \citep{sennrich2015neural} \footnote{To build the tokenizer we utilized SentencePiece: \url{https://github.com/google/sentencepiece}}. \pcyin{This sentence could be moved to implementation details.} \pcyin{I moved this sentence from Section 5}

\pcyin{TBD: Expreiment Setup Subsection.}

\alex{I need to come back and clean this up}

>>>>

% inserted by alex
\usepackage{amsmath,amssymb}
\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}
\DeclareMathOperator{\E}{\mathbb{E}}
\usepackage[dvipsnames]{xcolor}
\newcommand{\GH}{\textsc{Github}}
\usepackage{listings}
\usepackage{cleveref}

\lstdefinelanguage
  [x64]{Assembler}
  [x86masm]{Assembler}
  {basicstyle=\ttfamily\bfseries\scriptsize,
  frame=single,
  keepspaces=true,
  framesep=5pt,
  xleftmargin=10pt}
\lstset{language=[x64]{Assembler}}

% Pengcheng: common symbols
\newcommand\p{\ensuremath{{\mathrm{F}_u}}}
\newcommand\popt{\ensuremath{{\mathrm{F}_o}}}

@@

Program optimization is a classical problem in computer science that has existed for over 50 years \cite{mckeeman1965peephole, allen1971}, and most years, entire tracks at compiler conferences are dedicated to it. Most optimizing compilers make use of semantics-preserving heuristic-based optimizations. These optimizing transformations generally need to be written by experts for an individual compiler. These transformations are then applied to an intermediate representation of the code during the compilation process when the program is being transformed from high-level code into executable machine code.  \ed{I think there is a sentence missing here that explains most optimizing compilers use hand-written/heuristics-based/rule-based optimizations.} \alex{Good catch, I added 1-2 now}.  In an effort to automatically create optimized programs without human-defined heuristics, the research community has pioneered automated optimization methods, or ``superoptimizers," based on brute-force search, heuristic search, and satisfiability modulo theories (SMT) solvers. These superoptimizers have been shown to outperform compiler-based optimizations in some cases; however, in practice they are generally too costly to use for all code during the compilation process.

Deep reinforcement learning has shown promise in addressing challenging tasks that search-based techniques had once struggled to scale to \cite{silver2017mastering}. Similar methods grounded in machine learning could hold promise for speeding up automated program optimization by applying learned knowledge quickly, as opposed to running a costly search or solver algorithm at compile time. However, to our knowledge, current research in machine-learning-based program optimization has been generally constrained to the problem of expression simplification in a relatively simple domain-specific language (DSL) on synthetically created expressions and testcases \cite{shi2020}. Similarly, much work in machine-learning-based program synthesis  has been performed on synthetically-created datasets \cite{parisotto2016neuro,bunel2018leveraging}. While work on these datasets is an important research direction, these methods may have not yet been evaluated on realistic programs and input/output (I/O) examples, such as those found in open source code projects. Additionally, many of these machine-learning-based models often require access to a grammar and may not be directly extendable to program representations in-which a grammar may not be available or easy to construct.

\gn{Introducing a dataset doesn't sound exciting as discussing the core experimental result, which I think you should do first. Like ``we demonstrate that ML-based techniques can optimize real-world programs''} \alex{Restructured the following paragraph given the comments}

In this work, we demonstrate the ability of deep neural networks to optimize programs mined from real-world projects on \textsc{Github} using no other knowledge besides demonstration. We show that by performing supervised learning on heuristic-based compiler optimizations, the model is capable of learning non-trivial optimization strategies such as register allocation: a classic problem in compiler theory. Moreover, we also show that an iterative hill-climbing algorithm that has similarities to self-imitation learning \cite{oh2018self} is capable of discovering new optimizations. Some of the optimizations we witnessed the model discover include tricks to merge branches to shrink code, expand branches to reduce instructions executed, and a non-trivial ability to remove redundant instructions. In both the training and evaluation process we incorporate the use of SMT solvers, which can be used to represent programs in formulas on first-order logic, and show how the added rigor helps address challenges in synthesizing spurious programs. Based on our preliminary results, we argue that future machine-learning-based program synthesis work should strive to incorporate solvers and/or comprehensive testcase coverage for evaluation.  \pcyin{Maybe one or two more sentences to briefly discuss results?} \alex{Re-structured per Graham's comments + included more exciting results on top}

@@

The general undecidability of program equivalence means that there may always be room for improvement in optimizing programs~\citep{rice1953}. This is especially true as hardware options and performance goals become more diverse: what transformations are best for a scenario may vary greatly on performance objectives such as such as energy consumption or runtime or other factors. There are other classes of optimizations that are difficult or impossible to express in universal heuristics and semantics-preserving transformations, especially for program optimizations that might require domain-specific knowledge \citep{rinard2006}, are written by hand (e.g.,~OpenBLAS\footnote{https://www.openblas.net}) or adapted to specific hardware \citep{fftw}. Given these challenges and the time-consuming nature of implementing expert-written optimizations, there are advantages to developing methods that do not rely on expert-written heuristic-based optimizations. 
\ed{This language is a bit ambiguous.  Modern optimizing compilers are automated even though they are based on manually specified expert rules.} \alex{Changed this now }

There are classes of functions that are amenable to exhaustive search, where the optimal sequence of calculations is found \citep{massalin1987superoptimizer}. However, these are limited to very small sequences of machine instructions. State of the art methods of superoptimization either rely on search-based procedures \citep{schkufza2013stochastic}, or constraint-based methods \citep{sasnauskas2017}. However, these methods have difficulty scaling to larger problems, and as a result, typically do not meet the performance requirements of real development scenarios at compile time.

% \ed{Are these super-optimizers too?} \alex{Yes in the sense I believe both claim to be superoptimizers; perhaps ``no" if we are strict in defining a superoptimzier to find the most optimzal sequence. I am not sure how souper works (it uses SMT solvers I think) but I know that STOKE (Schkufza 2013) cannot guaranteed optimal. I think we can call these ``superoptimziers" as that does seem to be the norm in the literature. }

@@

\paragraph{Machine-Learning-Based Program Synthesis} Program synthesis is the task of automatically generating a program that is consistent with a certain specification. Recent work in machine-learning-based program synthesis has traditionally used I/O examples as a way to provide the program specification. In \cite{bunel2018leveraging} the authors perform experiments in the Karel programming language \cite{pattis1981karel} an educational programming language that manipulates the actions of a robot in a gridworld.
% The programming language contains control flow operations such as loops and conditionals. 
In these tasks, programs are synthetically generated by sampling from a domain-specific language (DSL). I/O examples are randomly generated, but with full-branch coverage \ed{Seems like a contradiction?}, meaning each individual conditional must be triggered by at least one  testcase; however, this does not imply comprehensive test coverage of all paths through program execution. While the authors RL experiments achieved highest performance with top-1 program synthesis accuracy; maximum-likelihood estimation (MLE) on target programs gradually outperformed RL as the beam-size increased to 50.

@@

% \gn{use ``citet'' for citations that are part of a sentence. Also in other places (like the future work)}
The task of program optimization considers transducing an (unoptimized) program $\p$ into an optimized one $\popt$ such that $\popt$ runs more efficiently or has smaller memory footprints.
Examples of program optimizations can range from removing a redundant operation to complex techniques for allocating the large number of program variables to a limited number of cpu registers.\pcyin{ALEX: can you provide a very simple and brief example of program optimization here?}. \alex{Done.}
In this paper we focus on program optimization at assembly level, where $\p$ and $\popt$ are programs written in X86 assembly language.
The standard toolkit for optimizing X86 assembly code is \texttt{GCC}, an optimizing compiler equipped with predefined set of optimization heuristics.

%In describing the task of neural program optimization,  we build on notation from the program synthesis formulation in \citet{bunel2018leveraging}. The task of program optimization through neural sequence models 
%\gn{Do we need ``through neural sequence models''? Be best to skip this and make it more broad in the problem formulation.} 
%is transducing an (unoptimized) program $\text{F}_u$ into an optimized one $\text{F}_o$, where $\text{F}_o$ is better than $\text{F}_u$ under some cost function $\mathbfcal{C}$, such as runtime, memory footprint or energy consumption.
%and equivalent under a distance metric $\mathbfcal{D}$. 
%Our goal is to learn a neural model  $f_{\theta}: \text{F}_u \mapsto \text{F}_o$. 
%\pcyin{Defer introducing $\mathcal{D}$ to the later paragraph where this notation is used.}

Because each program is a function that maps inputs to outputs, we will use $I$ to represent the hardware state prior to executing the program (i.e., input) and $O$ to represent the hardware state after executing the program (i.e., output).
Specifically, our goal is to learn a \gn{again, is ``neural'' necessary?} \alex{done} model $f_{\theta}: \text{F}_u \mapsto \text{F}_o$ such that a model-generated (optimized) program $\hat{F_o}$ attains lower cost under a cost function $\mathcal{C}(\cdot)$ evaluated on a suite of $K$ input-output test cases $\{ IO \}_{k=1}^K$:
%on our synthesized program $\hat{\text{F}}_o$,
%and observing $\mathcal{C}$ which may measure approximate runtime, energy consumption, or other goals. Furthermore, we may evaluate our distance function $\mathcal{D}$ either as equivalence under all testcases, or through verification by using an SMT solver. We consider a program optimized when: 
% cost function $\mathbfcal{C}$ which is able to evaluate desired program behavior, such as approximate runtime or energy consumption. For the transductive objective, the learned model's objective is then the following: 
\begin{equation}
    \label{eqn:optimizaiton_goal}
    \begin{split}
        \mathbfcal{C} \Big(\hat{\textrm{F}}_{o}; \{IO_k\}_{k=1...K} \Big)  \ 
        < \ 
        \mathbfcal{C} \Big(\textrm{F}_{u}; \{IO_k\}_{k=1...K}\Big) \\
         s.t. \;\; \mathbfcal{D}( \ 
                        \hat{\textrm{F}}_{o}, \textrm{F}_{u} ) \ 
                        = 0  % \;\; \
        %  \hat{\textrm{F}}_{opt}^i(I_k^i) = \ 
        %         O_k^i , \\
            % \forall  \; k \;  \in 1...K 
    \end{split}
\end{equation}

where $\mathcal{D}(\cdot) \in \{ 0, 1 \}$ is a distance metric that measures if the model optimized program $\hat{F}_o$ is functionally equivalent to the original one $\p$. 
As we explain in Section xxx, the cost function $C$ considers metrics like runtime, and could be computed by directly executing the optimized program on the test cases $\{ IO \}_{k=1}^K$.
For conciseness, we will abbreviate $\mathbfcal{C} (\hat{\textrm{F}}_{o}; \{IO_k\}_{k=1...K})$  as $\mathbfcal{C} (\hat{\textrm{F}}_{o})$.

In situations in which we desire to learn optimizations through demonstrations, we could collect ground-truth optimized programs $\popt$ from an optimizing compiler (e.g., \texttt{GCC}).
Our training set is therefore initialized with $N$ tuples of an I/O test suite $\{IO_k^i\}_{k=1^K}$, an unoptimized program $\p$, and the compiler-optimized one $\popt$:
%For example, as we elaborate in Section xxx, the C program compiler GCC could be configured to perform different levels of optimization.
%In situations in which we desire to learn optimizations through demonstrations, we will denote a ground-truth optimized program, for example, from an optimizing compiler, as $\textrm{F}_o$. In our later experiments, prior to training, we initialize a dataset $D_o$ of $N$ tuples of an I/O test suite, an unoptimized function, and compiler-optimized function as demonstration: 
\begin{equation} 
    \begin{split}
    \label{eqn:init_dataset}
        D_o = \
                \bigg\{
                    \Big( \
                        \{IO_k^i\}_{k=1...K}, \
                        \textrm{F}_{u}^i, \
                        \textrm{F}_{o}^i, \
                        % \textit{LiveOut}^i\
                    \Big) \
                \bigg\}_{i = 1...N} \\
        % s.t. \;\;   \textrm{F}_{opt}^i(I_k^i) = \ 
        %     \textrm{F}_{uopt}^i(I_k^i) = O_k^i  \\
        % \forall  \; k \;  \in 1...K \;\;  \ 
        %     \forall \; i \;  \in 1...N 
        % C(\textrm{F}_{uopt}^i; \{IO_k^i\}_{k=1...K})  \ 
        %     \leq \ 
        %     C(\textrm{F}_{uopt}^i; \{IO_k^i\}_{k=1...K}) 
    \end{split}
\end{equation}
As we later explain in \cref{sec:learning}, our learning algorithm is initialized with such  compiler-optimized programs as targets, but could discover more efficient rewrites by directly exploring the space of possible optimizations.
Before this, we first introduce our benchmark dataset of optimizing real-world programs in \cref{sec:benchmark}.
\pcyin{TODO: config \texttt{cleverref}}
%In \S\ref{sec:benchmark} we describe one concrete instantiation of this problem formulation on real-world programs, and in \S\ref{sec:learning} we describe our methodology for modeling and learning $f_\theta$ within this setup.

@@

\alex{1:12 PM Thursday: will circle back to this section}

@@

In order to create a dataset of functions whose re-writes could be executed and evaluated, we further mined a subset of 15,175 functions for training, 775 for validation, and 942 for testing. \ed{The jump from 1.77 million examples to ~17K deserves more of an explanation.} \gn{Right, and it's probably because only about 17k finished processing right? It'd be nice to change 1.77 million to the number of ones we actually \emph{tried} to process, and explain why some were filtered out.} We were able to create testcases through random generation using the \textsc{STOKE} toolkit.\footnote{\url{https://github.com/StanfordPL/stoke}}   In order to create conservative benchmarks \gn{Needs more explanation.}, we eliminated any functions in our dataset that caused segmentation faults either when testcases were executed on the original unoptimized or optimized function. \ed{Does this mean that if a function dereferences a pointer argument it will likely be excluded?} We further removed any functions where the compiler optimized program was not logically equivalent to the unoptimized program using an SMT solver and equivalent under all testcases. \ed{I think this also needs more explanation.  It makes it sound like the compiler is to blame, but that's not really the case.} To further avoid learning spurious optimizations, we required that our programs were not equivalent under testcases to three pathological spurious program examples. 
%\pyin{Could you explain the last sentence more?}
\ed{I also found the previous sentence confusing.}

@@

Our model $f_{\theta}$ is based on a standard Transformer-based encoder \citep{vaswani2017attention}, but with a LSTM-based decoder \citep{hochreiter1997long,chen-etal-2018-best} instead.
This was mainly done to speed-up prediction within the iterative training loop described in \S\ref{sec:hillclimbing}, as transformers suffer from quadratic decoding time in sequence length \citep{chen-etal-2018-best}.
\gn{More detail. What are your inputs and outputs. Do you do subword tokenization? Can you give an example of what the inputs and outputs look like?} \alex{Done in section 6.1 setup: model architecture, data processing (canonicalization and bpe) as well as configuration of performance + correcteness}

\gn{You have a pretty extensive description below on why you \emph{don't} do what you don't do, but only one sentence explaining what you actually do :)} \alex{I think maybe now there is some more implementation-specific detail in 6.1, do we think that is good now ? }
While it is often advantageous to leverage a grammar for code generation tasks \citep{parisotto2016neuro,yin-neubig-2017-syntactic}, we chose not to do so in this setting. A primary reason was for simplicity. An additional reason was access to large amounts of data for supervised learning, which may implicitly learn syntax. Previous work in machine-learning-based program synthesis has demonstrated that while a grammar may help greatly in low-data program synthesis regimes, it may do little to improve both MLE and RL-based experiments when extensive data is available for pre-training \cite{bunel2018leveraging}.

@@

\pcyin{When explaining a model/algorithm, it's important to first briefly describe the intuition behind each component/sub-section at the very beginning. Can we say something like the follows here: With the model pre-trained on compiler-optimized examples, we then aim to learn more efficient optimization strategies by directly minimizing the expected cost of model-optimized programs.}
Policy gradient algorithms seem like a natural fit for learning how to optimize programs; however, methods based on maximum-likelihood estimation, or imitation learning, provide advantages in terms of stability. Training large neural networks with policy gradient algorithms can be difficult to tune. These difficulties are often attributed to the gradient's high variance which is exacerbated by other issues in program synthesis: a general sparsity of rewards, credit assignment over long sequences, and brittleness of program semantics.

% \alex{I think this point here contradicts my earlier point about needing to sample from a grammar.... If a grammar really isn't that necessary after pre-training, then is brittle syntax really an issue?? It may be good to leave this out or accept either the nuance proposed / contradiction}
% \pcyin{Perhaps just removing ``brittleness of program syntax'' is good?}

@@

\begin{algorithm}
\label{alg:hill_climbing}
\begin{algorithmic}
\STATE Initialize model $\textbf{Model}$ parameters $\theta$ from pre-trained model \\
\STATE Initialize dataset of program function pairs and testcases: \\  
    $D = D_o =  \Big\{ \big( \
    \{IO_k^i\}_{k=1...K}, \
    \textrm{F}_{u}^i, \
    \textrm{F}_{o}^i, \
    % \textit{LiveOut}^i\
    \big) \ 
    \Big\}_{i=1...N}$ \\
 \WHILE{not converged}
  \STATE Sample a batch $B_{\mathrm{ex}}$ from $D_o$\\
  \FOR{ $\big( \
    \{IO_k^i\}_{k=1...K}, \
    \textrm{F}_{u}^i, \
    \textrm{F}_{o}^i, \
    % \textit{LiveOut}^i\
    \big)$  in $B_{\mathrm{ex}}$} 
    \STATE  sample $\hat{\textrm{F}}_{o}^i \; \sim \; f_{\theta}(\textrm{F}_{u}^i)$ \\
    \STATE calculate 
    $\mathbfcal{C}(\textrm{F}_{o}^i)$, 
    and $\mathbfcal{D} (\hat{\textrm{F}}_{o}^i, \textrm{F}_{o}^i)$ 
    \IF{$\mathbfcal{D} (\hat{\textrm{F}}_{o}^i, \textrm{F}_{o}^i) == 0$ and \ 
    $\mathbfcal{C}(\hat{\textrm{F}}_{o}^i) < \ 
    \mathbfcal{C}(\textrm{F}_{o}^i) $ }
        \STATE  Replace $\textrm{F}_{o}^i$ with sample $\hat{\textrm{F}}_{o}^i$ in $D$ \\
    \ENDIF
  \ENDFOR
  \STATE Sample a batch $B_{\mathrm{tr}}$ from $D$\\
  \STATE Update $\theta$ via supervised learning on $B_{\mathrm{tr}}$ from $D$ \\
 \ENDWHILE\caption{Hill-Climbing for Program Optimization}
\end{algorithmic}
\end{algorithm}

@@

\paragraph{Model and Training} In our experiments, for our model $f_{\theta}$, we utilized a 3-layer transformer encoder with embedding dimension of 512 with 8 attention-heads and a 2-layer LSTM decoder with a hidden size of 1024 using the attention mechanism from \citet{bahdanau2014neural}. We utilize the Adam optimizer \cite{kingma2014adam} with the inverse square root schedule from \citet{vaswani2017attention}. We pre-trained the model for 33K steps and subsequently performed hill-climbing for an additional 5K steps.

\paragraph{Data Processing} Using data out  order to process the $F_u$ and $F_o$, we need to perform extra canonicalization in order to filter out noise. x86-64 often uses \textsc{goto} instructions to jump to different parts of a binary: this is one way that control flow is implemented. We canonicalize numerical locations with ordinal locations: this fully preserves function semantics while removing noise from the prediction task. We additionally pre-process the assembly with byte-pair encoding \citep{sennrich2015neural} \footnote{To build the tokenizer we utilized SentencePiece: \url{https://github.com/google/sentencepiece}}. \pcyin{This sentence could be moved to implementation details.}
